---
title: 'American Brewing: IBUs and ABV'
author: Darrick Sturgeon
date: '2018-06-14'
slug: american-brewing-ibus-and-abv
categories:
  - data visualization
tags:
  - ggplot2
  - R
description: Desc
hacker_news_id: ''
lobsters_id: ''
meta_img: /images/image.jpg
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE, 
                      fig.width = 10.5, fig.height = 4, 
                      comment = NA, rows.print = 16)
library(tidyverse)
library(ggthemes)
library(colorblindr)
```

```{r mydata}
beer <- read_csv("/home/datys/Downloads/beers.csv")
```

```{r}
popular_beers <- beer %>% 
  filter(style %in% c("American IPA", "American Brown Ale", "American Porter", "American Pale Ale (APA)", "Witbier", "American Amber / Red Ale", "American Blond Ale")) %>%
  filter(!is.na(abv)) %>%
  filter(!is.na(ibu))
```

```{r}
# I want to include several types of beer here, and include them in the same plot somehow.

beerplot <-ggplot(popular_beers, aes(x = abv, y = ibu, color = style)) + 
  geom_point() +
  theme(axis.text.x = element_text(angle = 30, hjust = 1)) +
  ggtitle("ABV vs IBU in American IPAs") +
  theme_bw() +
  scale_color_colorblind()

beerplot
```

### Type of Graph

This is a scatter plot.

### The Data

Alcohol content and bitterness rating of various U.S. beers.  This is from a dataset hosted on kaggle.

### The Audience

Someone with some interest in american craft beer, probably at least knows what some of the types are.

### What I'm Trying to Show

bitterness and alcohol percentage are enough to cluster some types of beer.

### How to read and what to look for

Looking for trends of like colors, you can notice how many of them clump together.

### Presentation tips

It is better in these cases to do the extra work if possible.  Fitting a GMM to this data could give some interesting results.  The next challenge is plotting level sets of the gaussian distributions, which I don't yet know in R.

### Variations and Alternatives

bubble plots can give a third dimension to the data of a scatter plot by scaling individual points.

area scatter charts can reduce the amount of data plotted by grouping big clumps together, can be good for extremely large data sets.

A step above that is a heat map, which works well for large datasets, but can be trickier with multiple classes.

### How I created it

Simply plotting ABV vs IBU in a number of popular styles of beer.  The results were not too intriguing.  Unfortunately, the data itself is low dimensional and lacks any ratings or interesting features to compare against.
