---
title: 'Gapminder: Continental Life Expectancies'
author: Darrick Sturgeon
date: '2018-06-14'
slug: gapminder-continental-life-expectancies
categories:
  - data visualization
tags:
  - ggplot2
  - R
description: Desc
hacker_news_id: ''
lobsters_id: ''
meta_img: /images/image.jpg
---

Here we take a look at countries with low life expectancy across continents.

<!--display settings and libraries-->
```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE, 
                      fig.width = 10.5, fig.height = 4, 
                      comment = NA, rows.print = 16)
```

```{r, include=FALSE}
library(ggplot2)
library(tidyverse)
library(gapminder)
```




```{r lowlifeexpect}
gapm <- gapminder
gapm_lifeexp <- gapm %>%
  group_by(year, continent) %>%
  filter(lifeExp <= quantile(lifeExp, .25)) %>%
  arrange(continent, year, lifeExp)
```

```{r plot_lle}
gapm_lifeexp <- gapm_lifeexp %>%
  filter(continent != 'Oceania')
ggplot(gapm_lifeexp, aes(x = year, y = lifeExp, color = continent)) +
  geom_point(alpha = 0.2) +
  geom_smooth(method = 'lm', se = FALSE, alpha = 1) +
  scale_color_brewer(type = 'qual', palette = 2)
```

### Type of Graph

This is a scatter plot with linear fit lines.

### The Data

Life expectancy for many countries spanning the continents.  Here we look at the life expectancy over time of the lower quartile of countries.

### The Audience

Anyone might be interested in trends regarding life expectancy, so this could be targeted towards a general audience of readers who might click on a science article.

### What I'm Trying to Show

The continents are quite separable in terms of trend and magnitudes.

### How to read and what to look for

The lines of best fit do not intersect at all.

### Presentation tips

For time series data, it's important to highlight differences over time.  Here we can see that each trend is completely separate.

### Variations and Alternatives

Unlike before with the time series plot, this scatter plot has a range of data at each time point.  An alternative here would be box plots, in fact, they would show a lot more information about each quartile in addition to the lower one, and may illustrate more interesting trends  (it was not the challenge in this lab excersize, however).

### How I created it

I split colors by continent after cutting out the upper quartiles from the dataset.  Then fitted another smooth line as a linear model in this instance as opposed to the smooth curve fit from the gdp example.




